peter_models

### 2024-12-09 draft_finalreport.qmd
update draft and mdoels

```{r echo = FALSE}
# Libraries
library(tidyverse)  
library(pROC)
library(vcd)
library(car)
```

@@

```{r}
# Cleaning and recoding dataset_new
dataset_new <- dataset_new |> 
  mutate(
    # Education recoding
    educ = case_when(
      S13 == "(1) Grades 1-8" ~ "No college experience",
      S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
      S13 == "(3) High School graduate or GED" ~ "No college experience",
      S13 == "(4) Some college" ~ "Some college experience",
      S13 == "(5) Associates, 2-year degree" ~ "College graduate",
      S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
      S13 == "(7) Post-graduate degree" ~ "College graduate"
    ),
    # Sexual orientation recoding
    sex_orientation = case_when(
      S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
      S3 == "(02) Gay or lesbian" ~ "Queer",
      S3 == "(03) Bisexual" ~ "Queer",
      S3 == "(04) Something else (Specify)" ~ "Queer",
      S3 == "(88) Refused" ~ "Refused/Don't Know",
      S3 == "(99) I don" ~ "Refused/Don't Know"
    ),
    # Urban vs rural recoding
    rural = case_when(
      S14 == "(1) Large urban area" ~ "Urban/suburban",
      S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
      S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
      S14 == "(4) Small town or small city" ~ "Urban/suburban",
      S14 == "(5) Rural area" ~ "Rural"
    ),
    # Race recoding
    race_recoded = case_when(
      S2_RACE_PRIME == "(1) White" ~ "White",
      S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
      S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
      S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
      S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
      S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
      S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
      S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"
    ),
    # Age recoding
    age_recoded = case_when(
      S5_AGE == "(1) 16-17" ~ "16-17",
      S5_AGE == "(2) 18-29" ~ "18-29",
      S5_AGE == "(3) 30-39" ~ "30-39",
      S5_AGE == "(4) 40-49" ~ "40-49",
      S5_AGE == "(5) 50-59" ~ "50-59",
      S5_AGE == "(6) 60-69" ~ "60-69",
      S5_AGE == "(7) 70 +" ~ "70+"
    )
  ) %>%
  mutate(
    # Converting variables to factors
    race_recoded = factor(race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
    age_recoded = factor(age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
    educ = factor(educ, levels = c("No college experience", "Some college experience", "College graduate")),
    rural = factor(rural, levels = c("Rural", "Urban/suburban")),
    sex_orientation = factor(sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know"))
  )

@@

# Model Building

@@

## Model 2: stepwise regression
```{r echo=FALSE, results=FALSE}
# Perform stepwise selection
model_stepwise <- step(model, 
                       direction = "both",
                       trace = 0)
```

@@

## Model Diagnostics
```{r echo=FALSE, results=FALSE}
# Set seed for reproducibility
set.seed(123)

@@

```{r echo=FALSE, results=FALSE}
# Calculate the ROC curve
roc_curve <- roc(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
                 predicted_probs_stepwise, 
                 levels = c("Not important", "Important"))

# Print the AUC
auc_value <- auc(roc_curve)
cat("AUC:", auc_value, "\n")

# Plot the ROC curve
plot(roc_curve, col = "blue", main = "ROC Curve")
abline(a = 0, b = 1, col = "red", lty = 2)  # Add diagonal line for random guess

@@

@@

(Odds Ratio Plot)

@@

```{r echo=FALSE}
# Calculate proportions
proportions <- prop.table(table(dataset$Q1R5))
proportions
# Define custom labels
custom_labels <- c("Not important", "Important")

@@

ggsave("Proportion_Outcome_variable_RQ2.png", width = 20, height = 10)
```

@@

```{r echo = FALSE}
# Filter for significant variables (p-value < 0.05)
significant_data <- results_model_stepwise[results_model_stepwise$PValue < 0.05, ]

@@

# Save the plot with larger dimensions
ggsave("Significant_Variables_Plot_Enhanced.png", width = 20, height = 10)
```

>>>>

```{r echo = FALSE}
# Libraries
library(tidyverse)  
library(pROC)
library(vcd)
library(car)
library(MASS) # package for model
library(caret) # package for confusion matrix
library(janitor)
library(brant)

```{r echo = FALSE, results = FALSE}
file_path <- "//Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda"

load(file_path)          # Load the RDA file
ls()                     # List objects loaded into the environment
dataset <- da39096.0001  # Access the dataset

@@

```{r, echo=FALSE, results=FALSE}
# Cleaning and recoding dataset_new
dataset_new <- dataset_new |> 
  mutate(
    # Education recoding
    educ = case_when(
      S13 == "(1) Grades 1-8" ~ "No college experience",
      S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
      S13 == "(3) High School graduate or GED" ~ "No college experience",
      S13 == "(4) Some college" ~ "Some college experience",
      S13 == "(5) Associates, 2-year degree" ~ "College graduate",
      S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
      S13 == "(7) Post-graduate degree" ~ "College graduate"
    ),
    # Sexual orientation recoding
    sex_orientation = case_when(
      S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
      S3 == "(02) Gay or lesbian" ~ "Queer",
      S3 == "(03) Bisexual" ~ "Queer",
      S3 == "(04) Something else (Specify)" ~ "Queer",
      S3 == "(88) Refused" ~ "Refused/Don't Know",
      S3 == "(99) I don" ~ "Refused/Don't Know"
    ),
    # Urban vs rural recoding
    rural = case_when(
      S14 == "(1) Large urban area" ~ "Urban/suburban",
      S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
      S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
      S14 == "(4) Small town or small city" ~ "Urban/suburban",
      S14 == "(5) Rural area" ~ "Rural"
    ),
    # Race recoding
    race_recoded = case_when(
      S2_RACE_PRIME == "(1) White" ~ "White",
      S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
      S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
      S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
      S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
      S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
      S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
      S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"
    ),
    # Age recoding
    age_recoded = case_when(
      S5_AGE == "(1) 16-17" ~ "16-17",
      S5_AGE == "(2) 18-29" ~ "18-29",
      S5_AGE == "(3) 30-39" ~ "30-39",
      S5_AGE == "(4) 40-49" ~ "40-49",
      S5_AGE == "(5) 50-59" ~ "50-59",
      S5_AGE == "(6) 60-69" ~ "60-69",
      S5_AGE == "(7) 70 +" ~ "70+"
    )
  ) %>%
  mutate(
    # Converting variables to factors
    race_recoded = factor(race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
    age_recoded = factor(age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
    educ = factor(educ, levels = c("No college experience", "Some college experience", "College graduate")),
    rural = factor(rural, levels = c("Rural", "Urban/suburban")),
    sex_orientation = factor(sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know"))
  )

@@

@@

```{r echo=FALSE, results=FALSE}
# Perform stepwise selection
model_stepwise <- step(model, 
                       direction = "both",
                       trace = 0)
```

@@

```{r echo=FALSE, results=FALSE}
# Set seed for reproducibility
set.seed(123)

@@

```{r echo=FALSE, results=FALSE}
# Calculate the ROC curve
# roc_curve <- roc(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
#                  predicted_probs_stepwise, 
#                  levels = c("Not important", "Important"))
# 
# # Print the AUC
# auc_value <- auc(roc_curve)
# cat("AUC:", auc_value, "\n")
# 
# # Plot the ROC curve
# plot(roc_curve, col = "blue", main = "ROC Curve")
# abline(a = 0, b = 1, col = "red", lty = 2)  # Add diagonal line for random guess

@@

```{r, echo=FALSE, results=FALSE}
# data load and cleaning for ordinal regression model
required_cols_ord <- c("Q29", # Outcome Variable; interest in politics
                   "Q271", # Predictor Variable; importance of race to identity
                   "S2_RACE_PRIME", # Race
                   "S3", # Orientation
                   "S3B", # Gender
                   "S5_AGE", # Age Category
                   "S13", # Educational Attainment
                   "S14", # Community Type
                   "S7", # Nativity
                   "Q21", # Party ID
                   "S2_M", # Multiracial

                   # Religion-related variables
                   "Q58R1",            # Catholic
                   "Q58R2",            # Protestant
                   "Q58R3",            # Other Christian
                   "Q58R4",            # Jewish
                   "Q58R5",            # Muslim
                   "Q58R6",            # Hindu
                   "Q58R7",            # Buddhist
                   "Q58R8",            # Mormon/LDS
                   "Q58R9",            # Folk Religion
                   "Q58R10",           # Daoist
                   "Q58R11",           # Shamanism
                   "Q58R12",           # Ancestral Religion
                   "Q58R13",           # Spiritual but not religious
                   "Q58R14",           # Atheist
                   "Q58R15",            # Agnostic

                   # Variables capturing experiences with racism
                   "Q627", # Treated unfairly or discriminated against ,
                   "Q628", # Location of discrimination
                   "Q629R1" # Unfair treatment due to racial background/ethnicity
                   )

# Create new df
data_filtered <- dataset[, required_cols_ord]

# Convert the required columns to factors
data_filtered[required_cols] <- lapply(data_filtered[required_cols_ord], as.factor)

# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  data_filtered[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data_filtered[[paste0("Q58R", i)]]))
}

# Create a new Religion variable
data_filtered$Religion <- "None"
data_filtered$Religion[data_filtered$Q58R1_numeric == 1] <- "Catholic"
data_filtered$Religion[data_filtered$Q58R2_numeric == 1] <- "Protestant"
data_filtered$Religion[data_filtered$Q58R3_numeric == 1] <- "Other Christian"
data_filtered$Religion[data_filtered$Q58R4_numeric == 1] <- "Jewish"
data_filtered$Religion[data_filtered$Q58R5_numeric == 1] <- "Muslim"
data_filtered$Religion[data_filtered$Q58R6_numeric == 1] <- "Hindu"
data_filtered$Religion[data_filtered$Q58R7_numeric == 1] <- "Buddhist"
data_filtered$Religion[data_filtered$Q58R8_numeric == 1] <- "Mormon/LDS"
data_filtered$Religion[data_filtered$Q58R9_numeric == 1] <- "Folk Religion"
data_filtered$Religion[data_filtered$Q58R10_numeric == 1] <- "Daoist"
data_filtered$Religion[data_filtered$Q58R11_numeric == 1] <- "Shamanism"
data_filtered$Religion[data_filtered$Q58R12_numeric == 1] <- "Ancestral Religion"
data_filtered$Religion[data_filtered$Q58R13_numeric == 1] <- "Spiritual but not religious"
data_filtered$Religion[data_filtered$Q58R14_numeric == 1] <- "Atheist"
data_filtered$Religion[data_filtered$Q58R15_numeric == 1] <- "Agnostic"

# Convert Religion to a factor
data_filtered$Religion <- factor(data_filtered$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", "None"))

# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
redundant_cols <- c(
  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
)

# Remove redundant columns from the dataset
dataset_clean_ord <- data_filtered[, !(names(data_filtered) %in% redundant_cols)]

dataset_clean_ord <- dataset_clean_ord %>% mutate(educ = case_when(S13 == "(1) Grades 1-8" ~ "No college experience",
                                                           S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
                                                           S13 == "(3) High School graduate or GED" ~ "No college experience",
                                                           S13 == "(4) Some college" ~ "Some college experience",
                                                           S13 == "(5) Associates, 2-year degree" ~ "College graduate",
                                                           S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
                                                           S13 == "(7) Post-graduate degree" ~ "College graduate"),
                                          sex_orientation = case_when(S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
                                                                     S3 == "(02) Gay or lesbian" ~ "Queer",
                                                                     S3 == "(03) Bisexual" ~ "Queer",
                                                                     S3 == "(04) Something else (Specify)" ~ "Queer",
                                                                     S3 == "(88) Refused" ~ "Refused/Don't Know",
                                                                     S3 == "(99) I don" ~ "Refused/Don't Know"),
                                          rural = case_when(S14 == "(1) Large urban area" ~ "Urban/suburban",
                                                            S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
                                                            S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
                                                            S14 == "(4) Small town or small city" ~ "Urban/suburban",
                                                            S14 == "(5) Rural area" ~ "Rural"),
                                          white = case_when(S2_RACE_PRIME == "(1) White" ~ "White", 
                                                            S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Non-White",
                                                            S2_RACE_PRIME == "(3) Black or African American" ~ "Non-White",
                                                            S2_RACE_PRIME == "(4) Asian American" ~ "Non-White",
                                                            S2_RACE_PRIME == "(5) American Indian/Native American" ~ "Non-White",
                                                            S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Non-White",
                                                            S2_RACE_PRIME == "(7) Native Hawaiian" ~ "Non-White",
                                                            S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "Non-White"),
                                          race_recoded = case_when(S2_RACE_PRIME == "(1) White" ~ "White",
                                                                   S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
                                                                   S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
                                                                   S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
                                                                   S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
                                                                   S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
                                                                   S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
                                                                   S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"),
                                          age_recoded = case_when(S5_AGE == "(1) 16-17" ~ "16-17",
                                                                  S5_AGE == "(2) 18-29" ~ "18-29",
                                                                  S5_AGE == "(3) 30-39" ~ "30-39",
                                                                  S5_AGE == "(4) 40-49" ~ "40-49",
                                                                  S5_AGE == "(5) 50-59" ~ "50-59",
                                                                  S5_AGE == "(6) 60-69" ~ "60-69",
                                                                  S5_AGE == "(7) 70 +" ~ "70+"))

dataset_clean_ord <- dataset_clean_ord %>% mutate(race_recoded = factor(dataset_clean_ord$race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
                                         age_recoded = factor(dataset_clean_ord$age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
                                         educ = factor(dataset_clean_ord$educ, levels = c("No college experience", "Some college experience", "College graduate")),
                                         white = factor(dataset_clean_ord$white, levels = c("White", "Non-White")),
                                         rural = factor(dataset_clean_ord$rural, levels= c("Rural", "Urban/suburban")),
                                         sex_orientation = factor(dataset_clean_ord$sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know")))

```{r, echo=FALSE, results=FALSE}
# Exploratory data analysis
# # Create a summary table for all categorical variables
variables <- c("Q29", "Q271", "S2_RACE_PRIME", "S3", "S3B", "S5_AGE", "S13", "S14", "S7", "Q21", "S2_M", "Q627", "Q628", "Q629R1", "Religion", "educ", "sex_orientation", "rural", "white", "race_recoded", "age_recoded")

# # Chi-square tests for each variable
# chi_square_results <- lapply(variables, function(var) {
#   chisq.test(table(dataset_clean_ord[[var]], dataset_clean_ord$Q29))
# })
# 
# # Print chi-square results
# names(chi_square_results) <- variables
# print(chi_square_results)
# 
# tabyl(dataset_clean_ord$Q29)
# tabyl(dataset_clean_ord$Q271)

```{r, echo=FALSE, results=FALSE}
# ordinal regression model and assessment
# loading cleaned data
load("/Users/pdeguz01/Documents/git/Data/2020 CMPS/Output/2020cmps_clean.RData")
dataset_clean_ord <- dataset_clean

dataset_clean_ord$Q271_reordered <- factor(dataset_clean_ord$Q271,
                            levels = c("(5) Not at all important",
                                     "(4) Slightly important",
                                     "(3) Moderately important",
                                     "(2) Very important",
                                     "(1) Extremely important"))

ord_mod_interactions <- polr(Q29 ~ Q271_reordered * S3B +
                               race_recoded +
                  age_recoded +
            educ +  # Educational Attainment
            Q21, # Party affiliation
          data = dataset_clean_ord, Hess=TRUE)

summary(ord_mod_interactions)
# 
# 
# pvals_interact <- pnorm(-abs(summary(ord_mod_interactions)$coef[,"t value"]))*2
# ctable_interact <- cbind(summary(ord_mod_interactions)$coef,pvals)
# 
# ctable_interact
# 
# (exp_coefs_interact <- exp(cbind(OR=coef(ord_mod_interactions),confint(ord_mod_interactions))))
# 
# head(ord_mod_interactions$fitted.values)
# head(predict(ord_mod_interactions))
# 
# confusionMatrix(predict(ord_mod_interactions), dataset_clean_ord$Q29, mode="everything")
# 
# # brant test
# brant(ord_mod_interactions)
# 
# # generating predictions
# new_dataset <- data.frame(race_recoded=c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA"),
#                           Q271_reordered=c("(5) Not at all important", "(4) Slightly important", "(3) Moderately important", "(2) Very important", "(1) Extremely important", "(4) Slightly important", "(5) Not at all important"),
#                           age_recoded = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+", "18-29"),
#                           S3B = c("(1) Man", "(2) Woman","(3) Non-binary","(4) Something else (Specify)", "(1) Man", "(2) Woman", "(1) Man"),
#                           educ = c("No college experience", "Some college experience", "College graduate", "No college experience", "Some college experience", "College graduate", "College graduate"),
#                           Q21 = c("(1) Republican",  "(2) Democrat","(3) Independent", "(4) Other party", "(1) Republican",  "(2) Democrat",  "(3) Independent"))
# predict(ord_mod_interactions, new_dataset, type="probs")
```

@@

```{r, echo=FALSE, results=FALSE}
# odds_ratio_plot <- function(model) {
#   # Get odds ratios and CIs
#   or_data <- exp(cbind(OR=coef(model),confint(model)))
#   
#   # Calculate p-values
#   p_values <- pnorm(-abs(summary(model)$coefficients[,"t value"]))*2
#   
#   # Keep only significant variables
#   sig_indices <- p_values[names(coef(model))] < 0.05
#   or_data_sig <- or_data[sig_indices, ]
#   
#   ggplot() +
#     # Add vertical line at OR = 1
#     geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +
#     # Add points and error bars
#     geom_point(aes(x = or_data_sig[,1], 
#                    y = reorder(rownames(or_data_sig), or_data_sig[,1])), 
#               size = 4, color = "blue") +
#     geom_errorbarh(aes(xmin = or_data_sig[,2], 
#                        xmax = or_data_sig[,3],
#                        y = reorder(rownames(or_data_sig), or_data_sig[,1])),
#                    height = 0.3, color = "black") +
#     # Customize theme
#     theme_minimal() +
#     theme(
#       axis.text.y = element_text(size = 6, color = "black"),
#       axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
#       axis.title.x = element_text(size = 12, face = "bold"),
#       axis.title.y = element_text(size = 12, face = "bold"),
#       plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
#       plot.margin = margin(1, 1, 1, 1, "cm"),
#       legend.text = element_text(size = 8),
#       legend.title = element_text(size = 9)
#     ) +
#     # Add labels
#     labs(
#       title = "Significant Variables with Confidence Intervals (Odds Ratio)",
#       x = "Odds Ratio",
#       y = "Variables"
#     )
# }
# 
# # Generate the plot
# odds_ratio_plot(ord_mod_interactions)
# ggsave("RQ1_oddsratioplot.png", width = 20, height = 10)

![](RQ1_oddsratio.png)

@@

```{r echo=FALSE, results=FALSE}
# Calculate proportions
proportions <- prop.table(table(dataset$Q1R5))
proportions
# Define custom labels
custom_labels <- c("Not important", "Important")

@@

#ggsave("Proportion_Outcome_variable_RQ2.png", width = 20, height = 10)
```

@@

```{r echo = FALSE, results=FALSE}
# Filter for significant variables (p-value < 0.05)
significant_data <- results_model_stepwise[results_model_stepwise$PValue < 0.05, ]

@@

# Save the plot with larger dimensions
#ggsave("Significant_Variables_Plot_Enhanced.png", width = 20, height = 10)
```

![](Significant_Variables_Plot_Enhanced.png){width="20"}

### 2024-12-09 peter_models.qmd
update draft and mdoels

>>>>

```{r}
odds_ratio_plot <- function(model) {
  # Get odds ratios and CIs
  or_data <- exp(cbind(OR=coef(model),confint(model)))

  # Calculate p-values
  p_values <- pnorm(-abs(summary(model)$coefficients[,"t value"]))*2

  # Keep only significant variables
  sig_indices <- p_values[names(coef(model))] < 0.05
  or_data_sig <- or_data[sig_indices, ]

  ggplot() +
    # Add vertical line at OR = 1
    geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +
    # Add points and error bars
    geom_point(aes(x = or_data_sig[,1], 
                   y = reorder(rownames(or_data_sig), or_data_sig[,1])), 
              size = 4, color = "blue") +
    geom_errorbarh(aes(xmin = or_data_sig[,2], 
                       xmax = or_data_sig[,3],
                       y = reorder(rownames(or_data_sig), or_data_sig[,1])),
                   height = 0.3, color = "black") +
    # Customize theme
    theme_minimal() +
    theme(
      axis.text.y = element_text(size = 6, color = "black"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
      axis.title.x = element_text(size = 12, face = "bold"),
      axis.title.y = element_text(size = 12, face = "bold"),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.margin = margin(1, 1, 1, 1, "cm"),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 9)
    ) +
    # Add labels
    labs(
      title = "Significant Variables with Confidence Intervals (Odds Ratio)",
      x = "Odds Ratio",
      y = "Variables"
    )
}

# Generate the plot
odds_ratio_plot(ord_mod)
